Light inextensible strings under tension in the Schwarzschild geometry

### 2018-10-03 string_microarticle.tex
say explicitly where the PDF files are created
typo
add Flamm's paraboloid reference plus slight rewording to re-attain two page maximum

\noindent where units in which~$2GM=c=1$ are used, has been known for
over a century~\cite{schwarzschild1916} and makes a number of
counter-intuitive predictions.  Here I consider the spatial
component~$ds^2= dr^2/\left(1-1/r\right) + r^2d\phi^2$, confined to
equatorial points~($\theta=\pi/2$); we understand~$1\leq r<\infty$
throughout.  A circular string at constant~$r$ will, trivially, have
proper length~$2\pi r$.

@@

This differential equation does not appear to have a simple analytical
solution; I present some numerical results below.  It is
straightforward to show that the radius of curvature of a taut string
becomes infinite, and the string has a point of inflection,
if~$\left(r'\right)^2=2r\left(r-1\right)$.

\section{Results}
Solutions to equation~\ref{rdashdash} fall into two distinct classes:
those with $r>1$ at all points, and those which attain $r=1$, in which
case the string is tangential to the event horizon, in general.
Considering first cases with $r>1$, numerical methods show that if the
closest approach is greater than a critical value of~$\simeq
1.076109317(1)$, the string does not self-intersect
(Figure~\ref{closest_approach_non_self_intersecting}); if the closest
approach is less than this value, the then string crosses itself
at~$\phi=\pi/2$ (Figure~\ref{closest_approach_self_intersecting}).
The case where the string is tangent to the event horizon
at~$r=1,\phi=0$ is shown in Figure~\ref{fixed_EH_intersection}.

\begin{figure}[h!]
\centering
\includegraphics[width=70mm]{closest_approach_nonselfintersecting.pdf}
\caption{Non-self intersecting strings arranged by distance~$d$ of
  closest approach to the event horizon, occurring tangentially
  at~$\phi=0$}
\label{closest_approach_non_self_intersecting}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=70mm]{closest_approach_selfintersecting.pdf}
\caption{Strings arranged by closest approach to the event horizon:
 self-intersecting strings shown for $\phi\geq 0$}
\label{closest_approach_self_intersecting}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=70mm]{one_free_end_fixed_EH_intersection.pdf}
\caption{Strings tangential to event horizon}
\label{fixed_EH_intersection}
\end{figure}

@@

%%Vancouver style references.
\bibliographystyle{model3-num-names}
\bibliography{refs}

>>>>

\noindent where units in which~$2GM=c=1$ are used, has been known for
over a century~\cite{schwarzschild1916}.  Here I consider the spatial
component~$ds^2= dr^2/\left(1-1/r\right) + r^2d\phi^2$, confined to
equatorial points~($\theta=\pi/2$); we understand~$1\leq r<\infty$
throughout.  A circular string at constant~$r$ will, trivially, have
proper length~$2\pi r$.

@@

This differential equation does not appear to have a simple analytical
solution; solutions correspond to a taut string confined to Flamm's
paraboloid~\cite{flamm1916}.  It is straightforward to show that the
string has a point of inflection
if~$\left(r'\right)^2=2r\left(r-1\right)$.

\section{Results}
Solutions to equation~\ref{rdashdash} fall into two distinct classes:
those with $r>1$ at all points, and those which attain $r=1$, in which
case the string is tangential to the event horizon, in general.
Numerical methods show that if the closest approach is greater than a
critical value of~$\simeq 1.076109317(1)$, the string does not
self-intersect (Figure~\ref{closest_approach_non_self_intersecting});
if the closest approach is less than this value, the then string
crosses itself at~$\phi=\pi/2$
(Figure~\ref{closest_approach_self_intersecting}).  The case where the
string is tangent to the event horizon at~$r=1,\phi=0$ is shown in
Figure~\ref{fixed_EH_intersection}.

\begin{figure}[h!] % Created by "closest_approach.R"
\centering
\includegraphics[width=70mm]{closest_approach_nonselfintersecting.pdf}
\caption{Non-self intersecting strings arranged by distance~$d$ of
  closest approach to the event horizon, occurring tangentially
  at~$\phi=0$}
\label{closest_approach_non_self_intersecting}
\end{figure}

\begin{figure}[h!] % created by "closest_approach2.R"
\centering
\includegraphics[width=70mm]{closest_approach_selfintersecting.pdf}
\caption{Strings arranged by closest approach to the event horizon:
 self-intersecting strings shown for $\phi\geq 0$}
\label{closest_approach_self_intersecting}
\end{figure}

\begin{figure}[h!] % created by "one_free_end_fixed_EH_intersection.R"
\centering
\includegraphics[width=70mm]{one_free_end_fixed_EH_intersection.pdf}
\caption{Strings tangential to event horizon}
\label{fixed_EH_intersection}
\end{figure}

@@

%%Vancouver style references.
\bibliographystyle{model3-num-names}
\bibliography{stringrefs}
